The dynamics of risk perception in a Mediterranean agroecosystem \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2020-02-18 manuscript/manuscript.Rmd
preliminary .tex file

bibliography: bibliography.bib
bibstyle: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: rticles::springer_article
---

@@

tibble(risk = seq(0, 1, .1),
      Barley = expected_yield_barley(risk),
      Wheat = expected_yield_wheat(risk)) %>%
  gather(type, yield, Barley:Wheat) %>%
  ggplot(aes(risk, yield)) +
  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .5)+
  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .5) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = indiff, color = 'black', size = 3) +
  geom_text(data = indiff, label = indiff_label, nudge_x = 0.1, nudge_y = 0.1) +
  theme_classic() +
  #scale_color_manual(name = 'Crop', values = c('darkgoldenrod', 'lightgoldenrod')) +
  scale_color_manual(values = c('#e66101', '#5e3c99'), name = '') +
  labs(x = expression('Drought probability (' * theta * ')'), y = 'Expected yield')
```

@@

```{r, fig.height =2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period. The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields, given the payoffs defined in Table 1.'}
sim_past %>%
  filter(between(water_year, -9500, -5500)) %>%
  mutate(time_step = ceiling(water_year / 50),
         year = time_step * 50) %>%
  group_by(year) %>%
  summarise(risk = sum(drought) / 50) %>%
  filter(year != min(year)) %>%
  ggplot(aes(year, risk, fill = risk, color = risk)) +
  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
  geom_col()  +
  scale_fill_distiller(palette = 'PuOr') +
  scale_color_distiller(palette = 'PuOr')+
  theme_minimal() +
  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
```

@@

# Risk perception in a Mediterranean agroecosystem

@@

b <- sim1 %>%
  group_by(time, crop) %>%
  tally() %>%
ggplot(aes(time, n, fill = crop)) +
  geom_bar(width = 1, stat = 'identity', position = 'fill') +
  #scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
  #  scale_fill_manual(values = c('#e66101', '#5e3c99'), name = '') +
    scale_fill_manual(values = c('#f1a340', '#998ec3'), name = '') + 
  labs(x = 'Years', y = 'Wheat proportion') +
  theme_classic()

@@

```{reval=FALSE}
library(furrr)
plan(multisession)

sim2 <- tibble(rep = 1:500,
       droughts = replicate(500, make_droughts(), simplify = FALSE)) %>%
  mutate(results = future_map(droughts, run_sim)) %>%
  select(-droughts) %>%
  unnest(results)
```

>>>>

bibliography: bibliography.bib
bibstyle: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: rticles::springer_article
header-includes:
- \usepackage{float}
- \floatplacement{figure}{tb}
---

@@

tibble(risk = seq(0, 1, .1),
      Barley = expected_yield_barley(risk),
      Wheat = expected_yield_wheat(risk)) %>%
  gather(type, yield, Barley:Wheat) %>%
  ggplot(aes(risk, yield)) +
  geom_hline(yintercept = indiff$yield, linetype = 2, alpha = .5)+
  geom_vline(xintercept = indiff$risk, linetype = 2, alpha = .5) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = indiff, color = 'black', size = 3) +
  geom_text(data = indiff, label = indiff_label, nudge_x = 0.1, nudge_y = 0.1) +
  theme_classic() +
  #scale_color_manual(name = 'Crop', values = c('darkgoldenrod', 'lightgoldenrod')) +
  scale_color_manual(values = c('#e66101', '#5e3c99'), name = '') +
  labs(x = expression('Drought probability (' * theta * ')'), y = 'Expected yield (t/ha)')
```

@@

```{r, fig.height = 2, fig.width = 6, fig.cap = 'Annual risk of wheat crop failure due to drought, averaged by fifty-year period. The dashed line indicates the level of risk beyond which one would plant barley over wheat to maximize subjective expected yields, given the payoffs defined in Table 1.'}
sim_past %>%
  filter(between(water_year, -9500, -5500)) %>%
  mutate(time_step = ceiling(water_year / 50),
         year = time_step * 50) %>%
  group_by(year) %>%
  summarise(risk = sum(drought) / 50) %>%
  filter(year != min(year)) %>%
  ggplot(aes(year, risk, fill = risk, color = risk)) +
  geom_hline(yintercept = indiff$risk, color = 'black', linetype = 2, size = 1) +
  geom_col()  +
  scale_fill_distiller(palette = 'PuOr') +
  scale_color_distiller(palette = 'PuOr')+
  theme_minimal() +
  labs(x = 'Years BP', y = 'Drought risk')  + theme(legend.position = "none")
```

@@

@@

b <- sim1 %>%
  group_by(time, crop) %>%
  tally() %>%
ggplot(aes(time, n, fill = crop, color = crop)) +
  geom_bar(width = 1, stat = 'identity', position = 'fill') +
  #scale_fill_manual(values = c('darkgoldenrod', 'lightgoldenrod')) +
  #  scale_fill_manual(values = c('#e66101', '#5e3c99'), name = '') +
    scale_fill_manual(values = c('#f1a340', '#998ec3'), name = '') + 
    scale_color_manual(values = c('#f1a340', '#998ec3'), name = '') + 
    labs(x = 'Years', y = 'Wheat proportion') +
  theme_classic()

@@

### 2020-02-18 manuscript/manuscript.tex
preliminary .tex file

@@

    \titlerunning{Risk perception in a Mediterranean agroecosystem}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  What was the temporal pattern of climate and climate change during the
  Holocene in the eastern Mediterranean?
\item
  How well could crop diversification strategies cope with this pattern,
  given farmers' imperfect risk perception and decision-making?
\end{enumerate}

Questions: How would the earliest farmers have made decisions (dealt
with uncertainty?) about crop diversification? ie what were the
different algorithms? -- better understand the decision context What are
the consequences of theses different decisions (i.e.~how well did they
perform) at different points in the history of the mediterranean How
were farmers able to detect changes in the environment, rather than
unconsciously adapt to them? what are the different consequences we'd
expect for each THIS is the most important question because with it I
can frame the results around and individual's perception of droughts in
the neolithic, and the populaiton-level distribution of drought
perceptions at given times

@@

The basic decision-making problem facing a farmer seeking to diversify
their crops can be thought of as a game against nature. We can represent
this simplified decision context as a ``game'' in a game theorietic
context. cite luce raiffa 1989 for game against nature, and milnor 1952,
Agrawal1968, cassidy1971, gould 1963 The ``game'' in this context is the
farmer's decision of which crops to plant and in what proportions, given
uncertianty in the future ``state-of-nature'', and nature varies between
several possible states such as dry and wet years. payoff matrix goes
here in reality continuous, but cite behavioral econ for people often
intuitively solve an easier problem when faced with a complex real-world
situation The exact values in the payoff matrix here are less important
than the relative payoffs in each quadrant. We can think of this game as
being the culturally-inherited object. our question is how do
individuals make decisions here? so in this case the cultural norm would
be to mix wheat and barley, but the learning is in the exact ratio --
both individual and social learning influence perceptions of eindividual
vs social learning tucker 2007

Estimates of yield volume (t/ha) for each crop type in wet and dry years
were derived from isotopic studies of ancient wheat and barley samples
from archaeological sites in the Mediterranean (Araus, Slafer, and
Romagosa 1999) (Table 1).

\begin{table}
\centering
\caption{Estimated yields for ancient wheat and barley varieties derived from [@Slafer1999].}
\begin{tabular}{|l|l|l|}
\hline

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Expected wheat and barley yields under increasing drought risk
with the point of indifference highlighted.}
\end{figure}

If an individual can learn about their environment, such as the risk of
drought in any given year, then they can behave more rationally by
trying to maximize their subjective expected utility If the probability
distributin of nature's moves is known, the farmer can choose the crop
mix that simply maximizes the expected crop yields given the empirical
frequency distirubtion of nature's moves. In the language of game
theory, this strategy is knoan as a ``game of ficticious play'' against
nature. This is an effective risk-managing strategy, but like all
similar strategies that adapt to a specific pattern of variability, it
is vulnerable to changes in the pattern of variability (Janssen eta al
2007) This strategy works well when the environment is stationary --
nature plays from a fixed probability distirubtion, but is vulnerable to
environmental nonstationarity. That is, when the mean or higher order
moments of the rainfall distribution shift, playing a game of fictitious
play can backfire because you're too stuck in your ways while the world
changes around you. Thus it is important to understand the dynamics of
risk perception -- how subjective risks, and thus subjective expected
utility -- rise and fall in uncertain environments.

\hypertarget{sec:2}{%
\section{The dynamics of risk perception}\label{sec:2}}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-2.pdf}

\hypertarget{drought-risks-in-the-middle-holocene}{%
\section{Drought risks in the middle
Holocene}\label{drought-risks-in-the-middle-holocene}}

\hypertarget{climate-risks-and-risk-perception}{%
\section{Climate Risks and Risk
Perception}\label{climate-risks-and-risk-perception}}

Using the bias-corrected climate model output, I divided each model year
into dry years and wet years. A dry year was any year where less than
300mm of rain fell during the wet season (October-March), the threshold
below which wheat crops will generally fail ({\textbf{???}}), and a
wet/normal year was defined as any year above this threshold.

Given the modeled patterns of wet and dry years, drought risk for any
particular year was defined as the proportion of the previous 50 years
that were dry years

\begin{equation}
    P_{dry} = \frac{\sum_{n=t-1}^{t-50} precip_n < 300}{50},
\end{equation} where \(t\) is the current time step. The 50 year time
span was selected to approximate the accumulated observational knowledge
of an individual farmer and their immediate household. As a result, this
approach does not allow for accumulated social learning, although it
could easily be extended to do so in future studies.

\hypertarget{climate-modeling-and-bias-correction}{%
\subsection{Climate Modeling and Bias
Correction}\label{climate-modeling-and-bias-correction}}

Estimates of drought risk in the eastern Mediterranean during the the
past 10,000 years were derived from outputs of the TraCE-21k
paleoclimate simulation available on the National Center for Atmospheric
Research's Earth System Grid repository
(\url{https://www.earthsystemgrid.org}). TraCE-21K is a state-of-the-art
simulation that uses a coupled atmosphere-ocean general circulation
model (GCM) to recreate the transient response of the global climate
system to orbital parameters and greenhouse gasses over the past 22,000
years from the Last Glacial Maximum to the present (Figure 2) (He 2011).
It generates physically consistent spatiotemporal climate dynamics,
driven by current best estimates of climate forcings (e.g.~orbit,
greenhouse gasses, glacial meltwater flux). The model simulates these
dynamics on a six hourly timescale, and model outputs are archived at a
monthly resolution.

TraCE-21k simulation outputs for the past 10,000 years were
bias-corrected using the CDF-t method and observed precipitation from a
weather station in near the town of Salihli in western Turkey. This
location was selected due to its long (\textasciitilde{}70yr)
observational record, the representativeness of western Anatolian
climate for the greater eastern Mediterranean, and its proximity to
archaeological sites with comparative evidence of crop diversification
strategies.

Now incorporate realistic drought variability from simulations.

\begin{verbatim}
## Loading required namespace: ncdf4
\end{verbatim}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Annual risk of wheat crop failure due to drought, averaged by
fifty-year period}
\end{figure}

\hypertarget{the-model}{%
\section{The Model}\label{the-model}}

Run an experiment. Simulate a time series with varying drought
frequency. This is a stepwise nonstationary process

@@

Redo these plots so its a single person, one who was born in a drought
period then experiences an ameliorarion and another who experiences a
shift to drought conditions later in life? basically this section neds
to work more with individuals and explore that stuff

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-14-1.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-15-1.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-16-1.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-17-1.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

Note in these plots above and below we're summing all the wheat and
barley in the population, compared to below which shows the wheat
proportions of individuals
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-1.pdf}
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-19-2.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-1.pdf}
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-20-2.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-1.pdf}
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-21-2.pdf}

\begin{verbatim}
## Picking joint bandwidth of 0.0146
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-1.pdf}

\begin{verbatim}
## Picking joint bandwidth of 0.0204
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-22-2.pdf}

make this a function to calculate payoffs for all wheat or all barley
strategies as a comparison
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-24-1.pdf}

\section{Results and Discussion}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Slafer1999}{}%
Araus, J. L., G. A. Slafer, and I. Romagosa. 1999. ``Durum wheat and
barley yields in antiquity estimated from 13C discrimination of
archaeological grains: a case study from the Western Mediterranean
Basin.'' \emph{Australian Journal of Plant Physiology} 26 (4): 345.
\url{https://doi.org/10.1071/PP98141}.

\leavevmode\hypertarget{ref-Garnsey1989}{}%
Garnsey, Peter. 1989. \emph{Famine and Food Supply in the Graeco-Roman
World: Responses to Risk and Crisis}. Cambridge University Press.

@@

>>>>

\usepackage{float}
\floatplacement{figure}{tb}

@@

    \titlerunning{The dynamics of risk perception in a Mediterranean agroecosystem}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  How likely are droughts to occur each year in the eastern
  Mediterreanean, and how did these risks change over the Holocene?
\item
  How might Neolithic farmers have percieved these changing drought
  risks, and what were the consequences for farming communities with
  respect to their collective ability to manage risk?
\end{enumerate}

@@

The basic decision-making problem facing a farmer seeking to diversify
their crops can be thought of as a game against nature. We can represent
this simplified decision context as a ``game'' in a game theorietic
context (Luce raiffa 1989, milnor 1952, Agrawal1968, cassidy1971, gould
1963). The ``game'' in this context is the farmer's decision of which
crops to plant and in what proportions, given uncertianty in the future
``state-of-nature'', and nature varies between several possible states
such as dry and wet years (Table 1). In reality continuous, but people
often intuitively solve an easier problem when faced with a complex
real-world situation (cite behavioral econ or simon). The exact values
in the payoff matrix here are less important than the relative payoffs
in each quadrant. We can think of this game as being the
culturally-inherited object. Our question is how do individuals make
decisions here? So in this case the cultural norm would be to mix wheat
and barley, but the learning is in the exact ratio.

\begin{table}
\centering
\caption{Estimates of yield volume (t/ha) for ancient wheat and barley varieties derived from [@Slafer1999].}
\begin{tabular}{|l|l|l|}
\hline

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{Expected wheat and barley yields under increasing drought risk
with the point of indifference highlighted.}
\end{figure}

If an individual can learn about their environment, such as the risk of
drought in any given year, then they can behave more rationally by
trying to maximize their subjective expected utility. If the probability
distributin of nature's moves is known, the farmer can choose the crop
mix that simply maximizes the expected crop yields given the empirical
frequency distirubtion of nature's moves. In the language of game
theory, this strategy is known as a ``game of ficticious play'' against
nature. This is an effective risk-managing strategy, but like all
similar strategies that adapt to a specific pattern of variability, it
is vulnerable to changes in the pattern of variability (Janssen eta al
2007) This strategy works well when the environment is stationary --
nature plays from a fixed probability distirubtion, but is vulnerable to
environmental nonstationarity. That is, when the mean or higher order
moments of the rainfall distribution shift, playing a game of fictitious
play can backfire because you're too stuck in your ways while the world
changes around you. Thus it is important to understand the dynamics of
risk perception -- how subjective risks, and thus subjective expected
utility -- rise and fall in uncertain environments.

\hypertarget{sec:2}{%
\section{Holocene drought risks in the eastern
Mediterranean}\label{sec:2}}

Estimates of drought risk in the eastern Mediterranean during the the
past 10,000 years were derived from outputs of the TraCE-21k
paleoclimate simulation available on the National Center for Atmospheric
Research. TraCE-21K is a state-of-the-art simulation that uses a coupled
atmosphere-ocean general circulation model (GCM) to recreate the
transient response of the global climate system to changes in the
Earth's orbit and greenhouse gasses over the past 22,000 years from the
Last Glacial Maximum to the present (He 2011). It generates physically
consistent spatiotemporal climate dynamics, driven by current best
estimates of climate forcings (e.g.~orbit, greenhouse gasses, glacial
meltwater flux). The model simulates these dynamics on a six hourly
timescale, and model outputs are archived at a monthly resolution.

TraCE-21k simulation outputs for the past 10,000 years were extracted
from the grid cell overlaying the northern Levant and southeastern
Turkey. This location was selected due to its proximity to
archaeological sites with comparative evidence of crop diversification
strategies. This region includes some of the earliest evidence of crop
domestication, including sites like Jericho and Can Hasan.

Using the climate model output, I divided each model year into dry years
and wet years. A dry year was any year where less than 300mm of rain
fell during the wet season (October-May), the threshold below which
wheat crops will generally fail ({\textbf{???}}), and a wet/normal year
was defined as any year above this threshold. Given the modeled patterns
of wet and dry years, drought risk for any particular year was defined
as the proportion of the previous 50 years that were dry years. Using
the simulated paleoclimate, rather than modern-day weather station
observations or paleoclimate proxies, allows us to estimate not just the
first order statistics of the climate system (mean, variance of dry
years) but also the higher order dynamics like serial autocorrelation in
the perisistence of wet and dry years. This presents a much more
realistic picture of the inherent year-to-year uncertainty in the
climate system, and presents a challenge to simple risk-managing
strategies that assume climatic risks are fixed.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{Annual risk of wheat crop failure due to drought, averaged by
fifty-year period. The dashed line indicates the level of risk beyond
which one would plant barley over wheat to maximize subjective expected
yields, given the payoffs defined in Table 1.}
\end{figure}

\hypertarget{sec:3}{%
\section{The dynamics of risk perception}\label{sec:3}}

In order to manage risk, decision-makers must be perceive what exactly
that risk is in the local environment. An individual's perceived risk
entails more than simply the ``objective'' risk in the external
environment (Tucker et al. 2013). A person's perceived risk depends in
large part on their past experience, and how those experiences shape
their expectations for the future. Likewise, the social responses to
risk will refelct the distribution of perceived risks in a population,
and by extension the collective memory of past droughts will have
consequences for the population's survival and evolution.

We can capture the influences of risk perception computationally by
simulating a population of ``belief-based agents.'' These agents develop
beliefs about the risk of drought in their environment through personal
experience. Each agent's perceived drought risks will change their
subjective expected crop yields from planting different ratios of wheat
to barley.

\hypertarget{bayesian-brains}{%
\subsection{Bayesian brains}\label{bayesian-brains}}

We can efficiently simulate these population-level risk perceptions in a
population of agents with Bayesian brains (Kahvalti et al 2019 double
check). Bayesian learning allows for a minimal representation of the
environemnt, so that each agent's beliefs can be encoded in two
parameters -- defining the Beta distribution over their prior beliefs --
rather than the record of all observed years. Not only does this
approach have a clear computational efficiency, but for the same reason
this reflects the processes actually going on in the heads of decision
makers. Our brains do not record every bit of perceived informaiton in
memory, rather it stores a ``compact encoding'' of that information
which it uses for future decisions. Even if individuals aren't --
consciously or unconsciously --- making these calculations -- the basic
algorithmic problem faced by the brain, and solutions it has evolved,
reflect the same basic sets of constraints (Sanborn and Chater 2016).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Development of an individual's perceived drought risk with
time, assuming a fixed drought risk of 0.5. Beliefs are represented as
Beta probabiity distributions, and the increased certainty with age
reflects the varying effective sample size of the Beta prior.}
\end{figure}

\hypertarget{bayesian-learning-from-bernoulli-observations}{%
\subsubsection{Bayesian learning from bernoulli
observations}\label{bayesian-learning-from-bernoulli-observations}}

We can treat this problem elegantly using bayesian statistics using
online bayesian learning from bernoulli objersbations (Bissiri 2010).
Put simply, the probability of a drought occuring in any given year is
treated as a coin flip (i.e.~bernoulli distribution) with paramter theta
reprenting the risk of drought. Uncertainty in the value of theta can be
represented with a beta distribution, constrained between 0 and 1
(Figure XX). We can thus represent the exact information content of each
individual agent's subjective experience of droughts using the diffusion
of the prior belief. What we see here is how an indiviudal's perception
of the risk of drought may start off diffuse at first (here as an
uniformative flat prior in which all values of theta are equally likely.
As each year of experience is observed, the agent updates its beliefs
accordingly, balancing the information in this year's observation with
the cumulative weight of experience.

A simple ``fixed'' bayesian learner thus gets more certain with age. For
young agents with weak priors, the information of each new year can
strongly influence their beliefs. For older agents who have experienced
many more years their priors will be stronger, and they will be less
likely to update their beliefs when balancing the information from a
single year with the many decades of accumulated experience. As was the
case with the simple game of ficitious play above, this strategy is fine
when the environment is stable and risks do not change. But when there
is volatility in the environment, and risks can change, it becomes
important to change your mind. Being too inflexible in one's priors can
lead being too optimistic when things have really changed for the worse,
and alternately being too conservative when things finally do get
better. So from a Bayesian learning standpoint, how should agents choose
to believe new information.

\hypertarget{the-problem-of-nonstationarity}{%
\subsubsection{The problem of
nonstationarity}\label{the-problem-of-nonstationarity}}

That is, what if the risk of a drought, theta, is not a constant but
rather varies unpredictably in time? The statistical problem shifts then
to one of change detection (o'reilly 2013, gallistel et al 2014). In
other words do people simply adapt to changes -- as in the fixed
bayesian learner and moving average approaches -- or do they detect
them? Strict year-to-year updating can result in overly conservative
risk perceptions, suggesting people are slow to change their beliefs
when in reality they often quickly update them (Bissiri 2010). Indeed,
the ``forgetful bayesian brain'' model seems to perform better, with a
dynamic rather than fixed bayesian belief model (Zhang and Yu 2013).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Change in perceived drought risks in an older (50) and younger
(15) agent before and after a 25-year dry period.}
\end{figure}

So what are the consequences of these cognitive processes for early
farmers in the Mediterranean.

Simulate a time series with varying drought frequency. This is a
stepwise nonstationary process. Setup an experiment exploring 6
scenarios.

This suggests that people are more in danger of not realizing conditions
are improving when they actually have -- the opportunity cost -- rather
than failing to realize the change in the firstplace. The law of small
numbers is very important here. The range of each person's experience is
so small -- every individual's perceived risk will be a small and noisy
sample of the true value. Older people are much more certain, even
though they are wrong.

@@

@@

Note in these plots above and below we're summing all the wheat and
barley in the population, compared to below which shows the wheat
proportions of individuals.

\hypertarget{discussion-and-conclusions}{%
\section{Discussion and Conclusions}\label{discussion-and-conclusions}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Garnsey1989}{}%
Garnsey, Peter. 1989. \emph{Famine and Food Supply in the Graeco-Roman
World: Responses to Risk and Crisis}. Cambridge University Press.

@@

\leavevmode\hypertarget{ref-Tucker2013}{}%
Tucker, Bram, Jaovola Tombo, Tsiazonera, Patricia Hajasoa, Charlotte
Nagnisaha, Vorisoa Rene Lahitoka, and Christian Zahatsy. 2013. ``Beyond
Mean and Variance: Economic Risk Versus Perceived Risk of Farming,
Foraging, and Fishing Activities in Southwestern Madagascar.''
\emph{Human Ecology} 41 (3). Springer US: 393--407.
\url{https://doi.org/10.1007/s10745-013-9563-2}.
